Structured Versus Decorated Cospans

### 2022-03-16 ResponseToReviews.tex

%{\john It's hard to tweak the statement, since we're really trying to introduce the notation of decorations in maps of structured cospans here!    
%So 
%I've added something about this \emph{after} the statement of Theorem 2.1.  See if you like it:
%\begin{quote} 
%Note that the decoration $s \in F(m)$ is now an object in the category $F(m)$, not an element of a set as in Fong's original approach. 
%\end{quote}}

%{\chris Looks good!}

\item Proof of theorem 2.1: You should probably at least handwave that the cartesian liftings really are cartesian liftings. (No doubt it’s easy, but 
this 
is a classic sort of spot where errors can hide.)

\textbf{Kenny Courser carefully checked that these really are co/cartesian liftings, but the argument is straightforward and uninteresting so we will just 
add remarks saying that this can be shown.}   {\john $\checkmark$} {\chris $\checkmark$} {\kenny $\checkmark$}

\iffalse
{\kenny I think we originally had these details in our proof (of both Theorem 2.1 and Theorem 3.1) but removed them.}

{\john It's your evil advisor who removed those details.   I'll ``handwave it''---see above.}
\fi

@@

\item On page 9 during the construction of $\mathbb{H}$, I would appreciate a sentence giving some intuition for the functor $E$, at this point I’m confused about  its  role.

{\bf  We have added a remark on page 9 discussing this.}  {\john $\checkmark$} {\chris $\checkmark$} {\kenny $\checkmark$}

\iffalse
{\color{blue}{Kenny: We don't have any examples where $E$ is not an identity, but in my mind it's a functor that possibly modifies decorations, e.g. 
$F(a)$ is a category of graph structures on $a$, and $E(F(a))$ could be something like a category of possibly labeled graph structures on $a$, if 
that 
makes sense. Something potentially different than \emph{just} a graph structure on $a$.}}

{\chris Right! We can perhaps add a sentence to this end, saying that although we only use E as an identity, someone could need to change 
decorations 
for some application?}

{\kenny Sure, maybe sometime before the official statement of the Theorem when we're describing how the double functor $\mathbb{H}$ is defined?}

{\chris Sounds good to me!}

@@

{\bf Done.} {\chris $\checkmark$} {\kenny $\checkmark$} {\john $\checkmark$}

\item I think it would be worth saying in words a bit more prominently what the hypothesis of theorem 4.1 roughly means, namely that the fibers of 
the total 
category have finite colimits, reindexing preserves them, and the monoidal product on the fibers is cocartesian.

{\bf We have added a paragraph right below the statement of Theorem 4.1, clarifying its assumptions as suggested, which also incorporates an 
explanatory paragraph previously located under Corollary 4.3.} {\chris $\checkmark$} {\kenny $\checkmark$} {\john $\checkmark$}

\item I wonder if the “problem” where DCsp is not a structured cospan category could be “fixed” by upgrading the functor from Set to Cat in a 
non-discrete 
way, for example with an ordering with the zero vector field as the bottom element. I suppose this would give you different 2-cells in the resulting 
double category or bicategory, but perhaps it would yield the same 1-category.

{\bf This is an interesting idea but we haven't been able to get it to work yet.  See also comment 20.}

{\john $\checkmark$}
\end{enumerate}

### 2022-03-16 ResponseToReviews.tex

{\bf We have added a citation to Emily Riehl's book ``Category Theory in Context'' which includes said result, above Lemma 6.1.}

%{\john I have also changed ``well-known'' to ``known'', which is more polite to people who don't know it.} {\chris Good!}

{\john $\checkmark$} {\chris $\checkmark$} {\kenny $\checkmark$}

{\chris Third review starts here}
\item p. 2. ``Unfortunately, many applications of decorated cospans were flawed… for reasons painstakingly explained in [1].''

@@

Since fixing these problems is a major motivation for the whole paper, it would be good to explain what they are in slightly more detail, rather than 
just giving a reference. This could be done in a one paragraph or less. The gist of it, as I recall, is that in the original approach to decorated 
cospans, the isomorphism classes of open graphs depend on the identities of the edge elements, which is clearly wrong.

{\bf We have rewritten this paragraph to explain those flaws.}

@@

{\john $\checkmark$}

{\chris If I am not wrong, the paragraph cited here IS the explanation of the problem, perhaps in application terms rather than in theory terms. 
Consider perhaps adding one sentence along the lines suggested?}

{\kenny This seems similar to comment 14 above. So that's two people that want to know more about what the flaws are with the original decorated cospans but don't feel like opening our other paper...}

{\john You're right, Christina, I already explained why ``many applications of decorated cospans were flawed'' in this paragraph!   See if you like my boldface response.  I sneakily pretend that we rewrote this paragraph to explain the flaws.  We actually have rewritten this paragraph... but the explanation was already there!

And yeah, Kenny, it's annoying: people complain if you repeat already known stuff, but they also complain if you don't.}

@@

{\bf Done.}

{\chris $\checkmark$} {\john $\checkmark$}  {\kenny $\checkmark$}

\item A last general comment is that the paper unfortunately adopts the strategy, all too common in math papers, of putting all the examples in a 
section at 
the end of the paper. I think the readability of the paper would be improved, without much change to the organization, if the motivating example of 
open graphs were added to the relevant places in Sections 2 and 3. The remaining examples could be left in Section 6.

{\bf We believe that it is hard to analyze the examples (even the motivating one of open graphs) without having laid all relevant theory. Also, 
the comparison between the two formalisms---a major feature of Section 6---would then not be possible. In order to 
better direct the reader to the applications, we added one sentence in each introductory paragraph of Sections 2 and 3 that point to Section 6 for 
concrete examples, and also added a paragraph and the end of the Introduction (see comment 9). }

{\chris $\checkmark$} {\kenny $\checkmark$} {\john $\checkmark$}

@@

{\john
I don't want to bust up the examples and put some in sections 2 and 3. Anyone who wants examples can easily find them all in Section 6, we mention 
this section in the introduction (and perhaps should advertise it more there), and it's hard to analyze the examples without having all the theory in 
hand.}

{\chris I agree. How does the above boldface sound?}

{\kenny Agreed about keeping the examples together. The boldface looks good to me.}

{\john I tweaked the boldface a tiny bit, so you can revoke your checkmarks if you want; 
otherwise just comment out our little conversation here.}

### 2022-03-16 ResponseToReviews.tex

{\john He thought we were being dismissive because he didn't know that the ``tricks'' were all things I'd done myself!  If he looked at [1, Section 5] 
he'd see it was me coming up with a bunch of tricks to try to save the decorated cospan category of open graphs, and then Kenny and me explaining why 
this tricks don't work.  So maybe I should just delete ``despite all the tricks one might try''.  Sound good?}

{\chris Sounds good to me! See above, amend at will.}

{\kenny Sounds good.}

\item pg. 3: ``Many of the flawed applications of decorated cospans have been fixed using structured cospans.” This dismisses the work of other 
scholars 
using the decorated cospan approach as flawed. The alleged flaw should be made more precise. What about the decorated cospan approach is inferior to 
the structured cospan approach?''
%The goal of this paper should be to illustrate to an applied category theory audience the differences between 
%decorated and structured cospans. Referring back to previous work of the authors to make that point does not provide value.

{\bf The ``other scholars'' are John Baez, one of the authors here, and his coauthors.  The flaws---mathematical errors that render a number of claimed theorems incorrect---are now explained in  in the Introduction, in the paragraph beginning ``Unfortunately, many applications of decorated cospans were flawed.''  In the paper ``Structured cospans'', Kenny Courser and John Baez fixed all these mistakes using structured cospans.  See also comment (9).}

@@

{\chris Right, done. About our answer here, we could politely tell them that the problem is simply that they are wrong! And perhaps, in 
text amend that sentence to ``Many of the flawed applications of decorated cospans, \emph{like the open graph one described above},...'' because we 
do actually say what the flaw is there, correct? I don't see where we describe flaws in Section 6.}

{\kenny It looks like we loosely mention what the problem is on page 2, but we already explained (in detail) what the problems are in \emph{Structured cospans}. I don't see why we can't just point the reader that wants to see what those problems are to that paper. I don't think we explain any flaws in Section 6; only use both frameworks and compare them.}

{\kenny Actually, we do talk about the flaws a bit more in the middle of page 21 when we're setting up the example of open graphs for Theorem 2.2.}

{\john See if you like my boldface response.  I'm acting like we added an explanation of the flaws to our introduction, so we don't have to say ``Hey, buster --- we already explained this.''}

@@

And yeah, Kenny, it's annoying: people complain if you repeat already known stuff, but they also complain if you don't.}

\item p. 4, following statement of Thm 2.1.

The diagram depicting the decoration morphism $\tau$ as a natural transformation, first shown explictly on p. 10, should be moved here since it is 
implicitly used at the beginning of p. 5 to define the decoration of horizontal composites.

{\bf We have moved the natural isomorphism triangle as equation (1) in the statement of Theorem 2.1, and we have referred to it at page 10.} {\chris 
$\checkmark$} {\kenny $\checkmark$} {\john $\checkmark$}

@@

\item p. 9, ingredients for constructing a double functor of decorated cospan double categories.

Does the natural transformation $\theta \colon EF \Rightarrow F’ H$ have to be a natural isomorphism? Quite possibly I missed something, but I don’t see where in 
the following proofs that assumption is used.

{\bf We thank the referee for catching this important mistake!   It does not need to be an isomorphism, so we changed the construction in p. 
9 accordingly. In fact, in Section 6 we construct a double functor
\[      \graysquare \colon F \lCsp \to D \lCsp \]
sending any open Petri net with rates to its corresponding open dynamical system, and we do this
using a monoidal natural transformation $\theta$ that is \emph{not} an isomorphism.  We had mistakenly written that it was; we have fixed this now.}
{\john $\checkmark$} {\kenny $\checkmark$} {\chris $\checkmark$}

\iffalse
{\kenny I think they are correct in that we don't use this extra assumption.}

{\chris I have a vague feeling that we discussed something similar a long time ago, but I do not remember the context at all. Perhaps it was an 
%identity then? In any case, if we don't actually use it then we should make it into a transformation.}

{\kenny Yes, originally it was an identity. Structured cospans has similar squares for its maps, but those ones are required to be invertible. In my 
thesis, I was going to try to come up with a recipe for turning those squares into these ones, so I made these ones invertible, too, but never 
actually got around to this.}

{\chris I see! So for maps of structured cospans, invertibility is actually essential?}

{\kenny Yep, we needed certain maps $L' F \Rightarrow F' L$ (using structured cospan letters) to be invertible. We also built those maps by hand, 
though, and here, we're using Shulman's mystical machine to build these ones. I'm not sure what happens or is required when building maps of 
structured cospans using Shulman's machine...} {\chris Interesting!}

@@

\item p. 27, end of Section 6.

It seems that the left adjoint needed for structured cospans fails to exist because the chosen category of dynamical systems does not have enough 
morphisms. Relatedly, this is the only example where the pseudofunctor D: (FinSet,+) $\to$ (Cat, $\times$) comes from a simple functor D: (FinSet,+) 
$\to$ (Set, $\times$) by viewing sets as discrete categories. In the other examples, this would count as using decorated cospans in the old, flawed 
way. 
All of this suggests the question: is there a way of constructing a category of dynamical systems that has more morphisms, so that a connection with 
structured cospans can be established?

{\bf That's a great question.   We do not know how to construct such a category, but the referee
is correct in suggesting that doing it might give a structured cospan category of open dynamical
systems.

As the referee noticed, the ``old, flawed'' construction of 
decorated cospans from a lax monoidal functor $D \colon (\mathsf{FinSet},+) \to 
(\mathsf{Set}, \times) $ amounts to working with a discrete categories $D(S)$, treated
as sets.  Thus, our use of this construction amounts to working with a
set $D(S)$, rather than a full-fledged category, of dynamical systems on the finite set $S$.

In the example near the end of Section 6, the category $\int \! D$ can be seen as a category
of dynamical systems.   It is not discrete, but it does not have an initial object, as we'd
need to obtain a structured cospan
category that matches our decorated cospan category.

@@

{\chris This looks similar to comment 30.}

{\john Yes, but it's worth a separate answer.   I wrote an answer above.  See if you like it.  }

{\chris Sounds good! However I am a bit stuck on the following. Remember also in 
\href{https://thalis.math.upatras.gr/~cvasilak/documents/MonGroth.pdf}{MonoidalGrothendieck}, we had this issue of a lax monoidal pseudofunctor not 
really being a lax monoidal functor in the split case, whatever I mean here. This also arised for Brendan's graph if I am not wrong. Really what is 
going on is that for a (strictly) lax monoidal functor into Cat, the laxator is strictly natural and associativity is an equality, whereas a (weakly) 
lax monoidal functor into Cat may be strict as a functor, but the laxator is still pseudonatural and associativity still an isomorphism, as per our 
diagram (5). So I am not 100\% sure the starting statement in the second paragraph above is true. Let us think about that for a moment, it may end up 
being true in this case.}

{\john  Oh-ho!  In this response I don't want to get pulled into any subtle math questions that aren't  actually necessary.  I was just trying to give the referee a pat on the back.  See if you like my new simplified reply, and bless it with a check mark if you do.}

@@

{\john
I don't want to bust up the examples and put some in sections 2 and 3. Anyone who wants examples can easily find them all in Section 6, we mention 
this section in the introduction (and perhaps should advertise it more there), and it's hard to analyze the examples without having all the theory in 
hand.}

{\chris I agree. How does the above boldface sound?}

{\kenny Agreed about keeping the examples together. The boldface looks good to me.}

{\john I tweaked the boldface a tiny bit, so you can revoke your checkmarks if you want; 
otherwise just comment out our little conversation here. By the way, you can comment out
a big chunk of text just by putting $\backslash$iffalse in front of it and $\backslash$fi at the end.  This is a special case of a trick for making some text appear only if some condition is true!}

### 2022-03-16 ResponseToReviews.tex

\section{Our Answer}

\begin{enumerate}

\item As the only point against, I feel there is a growing number of “cospan” articles. The present work seems like it could serve as a comparison of 
all of 
these, but perhaps that comparison could be more explicit and detailed: what is it done in each one of these “cospan” articles? is there anything 
that 
is not already explained in this one? If some of the articles about decorated cospans were found to contain errors, is this the only construction of 
decorated cospans that solves these errors?

{\bf The present work reviews all known previous work on decorated and structured cospans in Section 6, and we now point this out in the Introduction.   As we mention, our previous paper ``Structured 
cospans'' fixed the errors in previous work using decorated cospans by using structured cospans instead.   Here we describe how the new improved 
decorated cospans can also fix those errors.     There may be even more ways to correct these errors, but we do not know them.}  {\kenny $\checkmark$} {\john $\checkmark$}

@@

{\chris As was suggested by John at some point, would we perhaps want to add a sentence or two in the introduction, saying that indeed there are a 
number of cospan articles etc. and that in the beginning of section 6 we elaborate on those?}

{\kenny I'm fine either way on this.}

{\john I have added a remark in the Introduction, ``This result [Theorem 4.1] shows that under certain conditions, structured and decorated cospans provide equivalent ways of describing open systems.  We illustrate this in Section 6 with applications to graphs, electrical circuits, Markov processes, Petri nets, Petri nets with rates, and dynamical systems.   This is meant to be a fairly thorough review of the existing literature.''

If you're happy with my response above give me a check mark, Christina!}

@@

{\chris Sounds good to me! See above, amend at will.}

{\kenny Sounds good.}

\item pg. 3: ``Many of the flawed applications of decorated cospans have been fixed using structured cospans.” This dismisses the work of other 
scholars 
using the decorated cospan approach as flawed. The alleged flaw should be made more precise. What about the decorated cospan approach is inferior to 
the structured cospan approach?''
%The goal of this paper should be to illustrate to an applied category theory audience the differences between 
%decorated and structured cospans. Referring back to previous work of the authors to make that point does not provide value.

{\bf The ``other scholars'' are John Baez, one of the authors here, and his coauthors.  The flaws---mathematical errors that render a number of claimed theorems incorrect---are now explained in  in the Introduction, in the paragraph beginning ``Unfortunately, many applications of decorated cospans were flawed.''  In the paper ``Structured cospans'', Kenny Courser and John Baez analyzed these mistakes in detail and fixed them using structured cospans.  See also comment (9).}

@@

{\kenny Actually, we do talk about the flaws a bit more in the middle of page 21 when we're setting up the example of open graphs for Theorem 2.2.}

{\john See if you like my boldface response.  I'm acting like we added an explanation of the flaws to our introduction, so we don't have to say ``Hey, buster --- we already explained this.''}

\item pg. 27: To show that U does not have a left adjoint, we use the following well-known result. For readers of the TAC paper previously cited, 
that might 
be a well known result that does not require a citation, but my understanding of the Compositionality audience assumed that the ACT work was targeted 
readers that are more familiar with the applications and thus cannot be assumed to know such results.

{\bf We have eliminated the term ``well-known'' and added a citation to Emily Riehl's book ``Category Theory in Context'' which includes said result, above Lemma 6.1.}

### 2022-03-16 ResponseToReviews.tex

\section{Our Answer}

\begin{enumerate}

\item As the only point against, I feel there is a growing number of “cospan” articles. The present work seems like it could serve as a 
comparison of 
all of 
these, but perhaps that comparison could be more explicit and detailed: what is it done in each one of these “cospan” articles? is there anything 
that 
is not already explained in this one? If some of the articles about decorated cospans were found to contain errors, is this the only construction of 
decorated cospans that solves these errors?

{\bf The present work reviews all known previous work on decorated and structured cospans in Section 6, and we now point this out in the end of 
the Introduction.   As we mention, our previous paper ``Structured 
cospans'' fixed the errors in previous work using decorated cospans by using structured cospans instead.   Here we describe how the new improved 
decorated cospans can also fix those errors.     There may be even more ways to correct these errors, but we do not know them.}  {\kenny $\checkmark$} 
{\john $\checkmark$} {\chris $\checkmark$}

%{\chris As was suggested by John at some point, would we perhaps want to add a sentence or two in the introduction, saying that indeed there are a 
%number of cospan articles etc. and that in the beginning of section 6 we elaborate on those?}

%{\kenny I'm fine either way on this.}

%{\john I have added a remark in the Introduction, ``This result [Theorem 4.1] shows that under certain conditions, structured and decorated cospans 
%provide equivalent ways of describing open systems.  We illustrate this in Section 6 with applications to graphs, electrical circuits, Markov 
%processes, Petri nets, Petri nets with rates, and dynamical systems.   This is meant to be a fairly thorough review of the existing literature.''

%If you're happy with my response above give me a check mark, Christina!}

{\chris Looks good, I only added ``in the end of'' the introduction, if it is OK, John, erase current comment.}

### 2022-03-16 ResponseToReviews.tex

 \item Lemma 2.4: when talking about the functors $F : (\mathsf{A},\otimes)\to(\mathsf{Cat},\times)$, I think sometimes it is specified that they are lax and sometimes 
it is not. Perhaps it would be clearer to always use “lax monoidal” or to just use “monoidal” to mean lax monoidal.

{\bf We have added ``lax'' at Lemma 2.4 in two occurences, and have checked the rest of the document to verify that this does not happen 
elsewhere.}  {\john $\checkmark$} {\chris $\checkmark$} {\kenny $\checkmark$}

\item Theorem 4.1: perhaps it would be clearer to say it “factors through R : Rex $\to$ SymMonCat”.

{\bf Done, also in Corollary 4.3.} {\chris $\checkmark$} {\kenny $\checkmark$}

@@

{\bf Done.} {\john $\checkmark$} {\chris $\checkmark$} {\kenny $\checkmark$}

\item COVID paragraph: I appreciate the effort into explaining the applied aspect. However, claims about the software (it is “easy”, we “easily”  modify, we “better understand”) sound too vague (is it computationally better? do we assume that the final user understands structured cospans? are we comparing  it to other Petri net editors?). Perhaps this paragraph could be fused as a comment to the previous one, which is already a very good justification  for the use of Petri nets with rates.

{\bf There is now a paper detailing how this software works, so according to this suggestion we have changed our paper so it simply refers to that other paper.}   {\john But we still need to add an arXiv link in reference 1!  It should be on the arXiv soon.}

{\kenny Can we put ``In preparation." if they don't have this on the arXiv in a timely manner?}

@@

{\kenny The boldface looks good to me, but I also like Christina's suggestions above if we really want to make an effort to appease this referee. I'm fine either way on this.}

{\john See if you like my new remarks in boldface, and give me a check mark if you do!  Kenny had checked this one off already, but I've changed it a lot.}

\item Given that Petri nets with structured cospans works, and Reaction Networks work, but open dynamical systems do not, I would like some more 
explanation of the differences here. The examples of graph, labeled 
graph, Petri net, reaction net are all examples of attributed C-Sets [1], the structured cospan construction in provided for arbitrary attributed 
C-Sets in Catlab.CategoricalAlgebra [2]. The open dynamical systems are not attributed C-Sets.(...)The examples in section 6 would be much stronger 
if the others included a theorem that all (or finitary) categories of presheaves satisfy the 
conditions of theorem 4.1 or a counter example (...) Hopefully it is easy to check and provide some 
commentary 
on it in this paper.

{\bf Unfortunately not all cases of C-sets meet the conditions of Theorem 4.1.   For example, the usual forgetful functor from $\mathsf{Graph}$ to $\mathsf{Set}$ sending any graph to its set of vertices is an opfibration, but not the forgetful functor sending any graph to its set of vertices {\kenny edges?}.  So, this gives a structured cospan category to which Theorem 4.1 does not apply.   Finding extra conditions such that it \emph{does} apply is an interesting question, but it would take more thought.}

{\john $\checkmark$} {\kenny $\checkmark$}

@@

\vskip 1em

This is a good enough situation to build a structured cospan category, but the right adjoint $f^\ast$ may not be an opfibration, in which case we can't apply Theorem 4.1 --- right?

With a lot of help from Todd Trimble's I once figured out when $f^\ast$ \emph{is} an opfibration, but remembering that argument and writing this up would take some work.  And I don't really feel like expanding the paper in such a way --- we've got more than enough already.   So how about if just tell the referee this is a great idea but it would take more thought?   See if you like my boldface comment above.  
}

{\kenny I think I recall you trying to explain this to me once: I believe the condition for $f^*$ to be an opfibration was that each hom-set $\hom(d,d)$ was a singleton, and that it was an if and only if. I remember trying to prove it, but getting lost in some of the details. Regardless, I agree that we have enough. The bold looks good to me.}

@@

{\chris Sounds good to me! See above, amend at will.}

{\kenny Sounds good.}

\item pg. 3: ``Many of the flawed applications of decorated cospans have been fixed using structured cospans.” This dismisses the work of other 
scholars 
using the decorated cospan approach as flawed. The alleged flaw should be made more precise. What about the decorated cospan approach is inferior to 
the structured cospan approach?''
%The goal of this paper should be to illustrate to an applied category theory audience the differences between 
%decorated and structured cospans. Referring back to previous work of the authors to make that point does not provide value.

{\bf The ``other scholars'' are John Baez, one of the authors here, and his coauthors.  The flaws---mathematical errors that render a number of claimed theorems incorrect---are now explained in  in the Introduction, in the paragraph beginning ``Unfortunately, many applications of decorated cospans were flawed.''.  In the paper ``Structured cospans'', Kenny Courser and John Baez analyzed these mistakes in detail and fixed them using structured cospans.  See also comment (9).}

{\john $\checkmark$} {\kenny $\checkmark$}

### 2022-03-16 ResponseToReviews.tex

{\chris Second review starts here}
\item Given the title and abstract of this paper “Structured vs Decorated Cospans” I would expect this paper to provide a judgment that structured  cospans are strictly better than decorated cospans in all circumstances and a justification for that claim, or a detailed analysis of the pros and cons of 
structured vs decorated cospans, some case studies in applications that show how the difference between these double categories is felt in the 
application domain, and a some description of the tradeoffs. This paper should come with some practical advice for the applied category theorist on 
how to choose whether to model their application with structured vs decorated cospans.

{\bf In fact our goal, as explained in the abstract and introduction, was to prove that the structured and decorated cospan approaches are equivalent under certain conditions, and to illustrate this result with a wide variety of examples.  These examples can serve as a guide for other similar applications, while new applications may require exploring which approach works best.  We do not make a stand that either is strictly better.   However, we have changed the end of the Introduction to read as follows:

\vskip 1em
\begin{quote}
This result shows that under certain conditions, structured and decorated cospans provide equivalent ways of describing open systems.  We illustrate this in Section 6 with applications to graphs, electrical circuits, Markov processes, Petri nets, Petri nets with rates, and dynamical systems.   This is meant to be a fairly thorough review of the existing literature.   It becomes clear that when either structured or decorated cospans can be used, structured cospans are simpler.  However, for open dynamical systems we apparently need decorated cospans, for reasons we explain.   Thus, to describe a map from open Petri nets with rates to open dynamical systems, we use \emph{both} structured and decorated cospans.
\end{quote}
}

{\john $\checkmark$} {\kenny $\checkmark$} {\chris $\checkmark$}
\iffalse
{\kenny I think the reader should come to their own conclusion about which framework they should use if they are trying to decide between the two, assuming both are applicable. I wouldn't say that one is better than the other. Personally, I would rather try structured cospans first, because I would rather think about a left adjoint than a symmetric lax monoidal pseudofunctor, but other people might feel differently.}

@@

{\kenny The boldface looks good to me, but I also like Christina's suggestions above if we really want to make an effort to appease this referee. I'm fine either way on this.}

\fi

{\john See if you like my new remarks in boldface, and give me a check mark if you do!  Kenny had checked this one off already, but I've changed it a 
lot.}

{\chris Looks great to me. Final two minor points, feel free to ignore or not and delete this comment. (1) Is ``apparently'' supposed to sound a bit 
sceptical? If not, would ``actually'' be better or not? (2) If we were being really accurate, don't we 
actually ONLY use decorated cospans to describe the map from open Petri nets with rates to open dynamical systems? Namely since dynamical systems only 
work as decorated cospans, we are required to use decorated cospans also for Petri nets to then get the map between them. Unless I am mixing something 
there.}

@@

{\kenny I think I recall you trying to explain this to me once: I believe the condition for $f^*$ to be an opfibration was that each hom-set $\hom(d,d)$ was a singleton, and that it was an if and only if. I remember trying to prove it, but getting lost in some of the details. Regardless, I agree that we have enough. The bold looks good to me.}

\item If we take the categorical systems theory perspective advanced by section 6, and the body of the authors’ recent work on functorial semantics 
for complex 
systems, then we should ask why the decorated and structured cospan double categories are isomorphic for reaction networks, but their semantics as 
dynamical systems are a decorated cospan category that is not equivalent to any structured cospan category. Is there a different construction of 
semantics for reaction networks that uses structured cospans?

{\bf We do not know a way to construct a (double) category of open dynamical systems that uses structured cospans; to find one would require getting around the result at the end of Section 6.  This is an interesting challenge.}

{\john $\checkmark$} {\kenny $\checkmark$} {\chris $\checkmark$}

\iffalse
{\chris Should we refer to which result we are talking about? Also, I am a bit confused about the question and answer. I thought reaction networks 
can be described in both ways?}

### 2022-03-16 ResponseToReviews.tex

\fi

{\john See if you like my new remarks in boldface, and give me a check mark if you do!  Kenny had checked this one off already, but I've changed it a 
lot.}

{\chris Looks great to me. Final two minor points, feel free to ignore or not and delete this comment. (1) Is ``apparently'' supposed to sound a bit 
sceptical? If not, would ``actually'' be better or not? (2) If we were being really accurate, don't we 
actually ONLY use decorated cospans to describe the map from open Petri nets with rates to open dynamical systems? Namely since dynamical systems only 
work as decorated cospans, we are required to use decorated cospans also for Petri nets to then get the map between them. Unless I am mixing something 
there.}

\item Given that Petri nets with structured cospans works, and Reaction Networks work, but open dynamical systems do not, I would like some more 
explanation of the differences here. The examples of graph, labeled 
graph, Petri net, reaction net are all examples of attributed C-Sets [1], the structured cospan construction in provided for arbitrary attributed 
C-Sets in Catlab.CategoricalAlgebra [2]. The open dynamical systems are not attributed C-Sets.(...)The examples in section 6 would be much stronger 
if the others included a theorem that all (or finitary) categories of presheaves satisfy the 
conditions of theorem 4.1 or a counter example (...) {\chris Without this analysis, I am left with the intuition
that for discrete structures like finitary presheaves over a finitely presented schema, the decorated and structured cospans are going
to be equivalent (...) If that intuition is correct, the paper should say something about it. If the intuition is incorrect, then additional 
explanation should be provided.} Hopefully it is easy to check and provide some commentary on it in this paper.

### 2022-03-16 ResponseToReviews.tex

{\chris I see, thanks!}
\fi

\item The development of the AlgebraicPetri software system is credited to Halter and Patterson based on a blog post. The correct citation is to the 
software itself and should be credited to the entire AlgebraicPetri team based on a citation to the repository [3]. The additional authors would be 
Andrew Baas and James Fairbanks.

{\bf We have added a reference to the new paper ``An
algebraic framework for structured epidemic modeling'' by A.\ Baas, J.\ Fairbanks, M.\ Halter, S.\ Libkind and E.\ Patterson.  We have also added a reference to the AlgebraicPetri GitHub repository.}

{\john \checkmark} {\kenny \checkmark} {\chris \checkmark}

\iffalse
{\john It'd be good to add a reference to the repository.}

@@

{\chris I added the reference but looks a bit funny, edit at will. Did not add any text.}

{\john I've improved the reference and added a citation to it.  Is this okay now?}

{\chris Yes!}
\fi

%[Several points in the text are written in a way style that can be interpreted as the dismissive style common to many discussions of Category 
%Theory. I would recommend rewriting at least the following points that should be made in a more collegial tone.]

@@

\item pg. 2: “But this hope is doomed for reasons painstakingly explained in [1, Section 5] … despite all the tricks one might try.” This paragraph 
provides 
no insight into the reasons that those tricks fail. Referring them to tricks is dismissive. The citation is to a paper in TAC by the current authors.

{\bf We removed the part of the sentence ``despite all the tricks one might try'' according to this suggestion. In [3, Section 5], these attempts are 
explained in detail, and we find them of value because the diligent reader may in fact come up with some of these for fixing the problem.} {\chris 
\checkmark}

@@

{\kenny Sounds good.}

\item pg. 3: ``Many of the flawed applications of decorated cospans have been fixed using structured cospans.” This dismisses the work of other 
scholars 
using the decorated cospan approach as flawed. The alleged flaw should be made more precise. What about the decorated cospan approach is inferior to 
the structured cospan approach?''
%The goal of this paper should be to illustrate to an applied category theory audience the differences between 
%decorated and structured cospans. Referring back to previous work of the authors to make that point does not provide value.

{\bf The ``other scholars'' are John Baez, one of the authors here, and his coauthors.  The flaws---mathematical errors that render a number of claimed theorems incorrect---are now explained in  in the Introduction, in the paragraph beginning ``Unfortunately, many applications of decorated cospans were flawed.''.  In the paper ``Structured cospans'', Kenny Courser and John Baez analyzed these mistakes in detail and fixed them using structured cospans.  See also comment (9).}

{\john $\checkmark$} {\kenny $\checkmark$} {\chris \checkmark}

\iffalse
{\john The "other scholars" are mainly me and Brendan, and the flaws are that some claims are false! I think we describe the flaws in Section 6; if so 
we should point to that.}

@@

{\kenny It looks like we loosely mention what the problem is on page 2, but we already explained (in detail) what the problems are in \emph{Structured cospans}. I don't see why we can't just point the reader that wants to see what those problems are to that paper. I don't think we explain any flaws in Section 6; only use both frameworks and compare them.}

{\kenny Actually, we do talk about the flaws a bit more in the middle of page 21 when we're setting up the example of open graphs for Theorem 2.2.}

{\john See if you like my boldface response.  I'm acting like we added an explanation of the flaws to our introduction, so we don't have to say ``Hey, 
buster --- we already explained this.''}

\fi
{\chris I would potentially skip the first sentence of our response above. Does it sound a bit hostile? And perhaps he also means other people that 
may have used this approach, or perhaps Brendan himself (not a coauthor at the time?) that invented them. Not too important in any case.}

@@

{\chris Third review starts here}
\item p. 2. ``Unfortunately, many applications of decorated cospans were flawed… for reasons painstakingly explained in [1].''

Since fixing these problems is a major motivation for the whole paper, it would be good to explain what they are in slightly more detail, rather than 
just giving a reference. This could be done in a one paragraph or less. The gist of it, as I recall, is that in the original approach to decorated 
cospans, the isomorphism classes of open graphs depend on the identities of the edge elements, which is clearly wrong.

{\bf We have rewritten this paragraph to explain those flaws.}

\iffalse
{\chris If I am not wrong, the paragraph cited here IS the explanation of the problem, perhaps in application terms rather than in theory terms. 
Consider perhaps adding one sentence along the lines suggested?}

### 2022-03-17 cospans_new.tex
did a lot of work, almost done now

% Structured versus decorated cospans
% John Baez, Kenny Courser and Christina Vasilakopoulou
% in LaTeX
% 2021/03/15 - dealing with referee's comments

@@

This result shows that under certain conditions, structured and decorated cospans provide equivalent ways of describing open systems.  We illustrate this in Section \ref{Applications} with applications to graphs, electrical circuits, Markov processes, Petri nets, Petri nets with rates, and dynamical systems.   This is meant to be a fairly thorough review of the existing literature.   It becomes clear that when either structured or decorated cospans can be used, structured cospans are simpler.  However, for open dynamical systems we apparently need decorated cospans, for reasons we explain.   Thus, to describe a map from open Petri nets with rates to open dynamical systems, we use \emph{both} structured and decorated cospans.

>>>>

% Structured versus decorated cospans
% John Baez, Kenny Courser and Christina Vasilakopoulou
% in LaTeX
% 2021/03/16 - dealing with referee's comments

@@

This result shows that under certain conditions, structured and decorated cospans provide equivalent ways of describing open systems.  We illustrate this in Section \ref{Applications} with applications to graphs, electrical circuits, Markov processes, Petri nets, Petri nets with rates, and dynamical systems.   This is meant to be a fairly thorough review of the existing literature.   It becomes clear that when either structured or decorated cospans can be used, structured cospans are simpler.  However, for open dynamical systems we need decorated cospans, for reasons we explain.   Thus, to describe a map from open Petri nets with rates to open dynamical systems, we use decorated cospans.
